A generalization of the Bradley-Terry model for draws in chess

### 2019-05-16 inst/collusion_chess_draws.Rnw
work

\begin{figure}[htbp]
\begin{center}
<<ddcc,echo=TRUE,fig=TRUE>>=
pie(maxp(kka_3draws))
@
\caption{Maximum likelihood estimate for the strengths of the 3 Grandmasters in Table~\ref{kka} 
under a likelihood function that admits differential proclivities to draw}
  \end{center}
\end{figure}

@@

\section{Chess collusion: Cura\c{c}ao 1962}

The method of support allows one to account for proclivity to draw
while accounting for players' strengths and first-move advantage.  We
now consider one famous incident, the World Chess Championships 1963,
in which Fischer famously and publicly claimed that the Soviet players
colluded.  The accusation was that Soviet players agreed ahead of time
to draw when playing one another, in order to conserve energy for
subsequent matches~\citep{moul2009}.

@@

% \begin{table}
% \centering
% \begin{tabular}{ll|cccccccc}
% &
% &%\rotatebox{-90}{Petrosian} 
% &\rotatebox{-90}{Keres} 
% &\rotatebox{-90}{Geller}
% &\rotatebox{-90}{Fischer} 
% &\rotatebox{-90}{Korchnoi}
% &\rotatebox{-90}{Benko}
% &\rotatebox{-90}{Tal}
% &\rotatebox{-90}{Filip}\\
% Petrosian &USSR    &             &DDDD    &DDDD   &DWDD   &DDWW   &DDWD   &WWD    &DWWD\\
% Keres     &USSR    &             &NA      &DDDD   &LDWD   &DDWD   &WWWL   &WDW    &DWWD\\
% Geller    &USSR    &             &NA      &NA     &WWDL   &DDWD   &DDDW   &DWW    &DWWD\\
% Fischer   &USA     &             &NA      &NA     &NA     &LWLD   &LWDW   &DWD    &WDWD\\
% Korchnoi  &USSR    &             &NA      &NA     &NA     &NA     &DDDL   &WLD    &WWWW\\
% Benko     &USA     &             &NA      &NA     &NA     &NA     &NA     &WLD    &LWWD\\
% Tal       &USSR    &             &NA      &NA     &NA     &NA     &NA     &NA     &WLD \\
% Filip     &TCH     &             &NA      &NA     &NA     &NA     &NA     &NA     &NA \\
% \end{tabular}
% \caption{Cura\c{c}ao 1962 candidates, world championship 1963}
% \end{table}

\section{Chess collusion: interzonal Stockholm 1962}

>>>>

%\begin{figure}[htbp]
%\begin{center}
%<<ddcc,echo=TRUE,fig=TRUE>>=
%pie(maxp(kka_3draws))
%@
%\caption{Maximum likelihood estimate for the strengths of the 3 Grandmasters in Table~\ref{kka} 
%under a likelihood function that admits differential proclivities to draw}
%  \end{center}
%\end{figure}

@@

\section{Collusion in the World Chess Championship 1963}

The method of support allows one to account for proclivity to draw
while accounting for players' strengths and first-move advantage.  We
now consider one famous incident, following the World Chess
Championships 1963, in which Fischer famously and publicly claimed
that the Soviet players colluded.  The accusation was that Soviet
players agreed ahead of time to draw when playing one another, in
order to conserve energy for subsequent matches~\citep{moul2009}.

Here I analyse the original subject of the complaint, the Cura\c{c}ao
1962 Candidates, Table~\ref{candidates}.  Fischer alleged that three
of the Soviet players (Keres, Petrosian, and Geller) had agreed to
draw amongst themselves.

@@

 \begin{table}
 \centering
 \begin{tabular}{ll|cccccccc}
 &
 &%\rotatebox{-90}{Petrosian} 
 &\rotatebox{-90}{Keres} 
 &\rotatebox{-90}{Geller}
 &\rotatebox{-90}{Fischer} 
 &\rotatebox{-90}{Korchnoi}
 &\rotatebox{-90}{Benko}
 &\rotatebox{-90}{Tal}
 &\rotatebox{-90}{Filip}\\
 Petrosian &USSR    &             &DDDD    &DDDD   &DWDD   &DDWW   &DDWD   &WWD    &DWWD\\
 Keres     &USSR    &             & -      &DDDD   &LDWD   &DDWD   &WWWL   &WDW    &DWWD\\
 Geller    &USSR    &             & -      & -     &WWDL   &DDWD   &DDDW   &DWW    &DWWD\\
 Fischer   &USA     &             & -      & -     & -     &LWLD   &LWDW   &DWD    &WDWD\\
 Korchnoi  &USSR    &             & -      & -     & -     & -     &DDDL   &WLD    &WWWW\\
 Benko     &USA     &             & -      & -     & -     & -     & -     &WLD    &LWWD\\
 Tal       &USSR    &             & -      & -     & -     & -     & -     & -     &WLD \\
 Filip     &TCH     &             & -      & -     & -     & -     & -     & -     & - \\
 \end{tabular}
 \caption{Cura\c{c}ao 1962 \label{candidates} Candidates, World Chess Championship 1963}
 \end{table}

\subsection{Chess collusion: interzonal Stockholm 1962}
